---
title: 'Compare HS Stats'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.width = 8, fig.height = 6)

library(tidyverse)
library(leaflet)
library(sf)
```

```{r}
data_dir <- file.path('..', 'data')
results_dir <- file.path('..', 'results')

pubprivhs_univ_df <- readRDS(file.path(data_dir, 'pubprivhs_univ_df.RDS'))  # created in create_cb_geo_hs_visits.R
hs_geomarket_distance_df <- readRDS(file.path(data_dir, 'hs_geomarket_distance_df.RDS')) %>% as_tibble()
```

```{r}
hs_df <- pubprivhs_univ_df %>% 
  filter(univ_id == 'all') %>% 
  select(hs_ncessch, hs_pct_free_reduced_lunch, hs_zip_inc_house_med)

plot_eps <- function(borders, y_var, hs_control = NULL) {
  hs_geomarket_distance_df %>% 
    filter(border %in% borders) %>% 
    mutate(
        eps = if_else(hs_eps == str_split(border, '-', simplify = T)[,1], 'EPS 1', 'EPS 2'),
        distance_category = case_when(
          distance_mi <= 0.5 ~ '<=0.5',
          distance_mi <= 1 ~ '<=1',
          distance_mi <= 2 ~ '<=2',
          .default = 'unrestricted'
        ) 
      ) %>% 
    inner_join(hs_df %>% filter(!is.na(get(y_var))), by = 'hs_ncessch') %>% 
    ggplot(aes(x = distance_mi, y = .data[[y_var]], shape = eps, color = distance_category)) + 
    geom_point() +
    scale_shape_manual(values = c('EPS 1' = 15, 'EPS 2' = 1)) +
    scale_color_manual(values = c('<=0.5' = 'red', '<=1' = 'darkorange', '<=2' = 'gold', 'unrestricted' = 'gray')) + 
    guides(
      shape = guide_legend(order = 1),
      color = guide_legend(order = 2)
    ) +
    theme_minimal()
}
```

```{r}
get_diff <- function(data_df, y_var) {
  agg_df <- data_df %>% 
    group_by(hs_eps) %>% 
    summarise(m = mean(get(y_var), na.rm = T), .groups = 'drop')
  
  if (nrow(agg_df) == 2) {
    return(abs(diff(agg_df$m)))
  }
}

get_std <- function(data_df, y_var) {
  sd(data_df[[y_var]], na.rm = T)
}

gen_agg_data <- function(agg_type, y_var) {
  data_df <- data.frame(
    border = character(),
    unrestricted = double(),
    less_than_2 = double(),
    less_than_1.5 = double(),
    less_than_1 = double(),
    stringsAsFactors = F
  )
  
  if (agg_type == 'diff') f <- get_diff
  if (agg_type == 'std') f <- get_std
  
  # For each border
  for (b in unique(hs_geomarket_distance_df$border)) {
    # Filter for all HS on either side of border
    x <- hs_geomarket_distance_df %>% filter(border == b) %>% inner_join(hs_df, by = 'hs_ncessch')
    
    # Calculate difference in mean for no border restriction, <=2 mi, <= 1.5 mi, and <=1 mi
    data_df <- data_df %>% add_row(
      border = b,
      unrestricted = f(x, y_var),
      less_than_2 = f(x %>% filter(distance_mi <= 2), y_var),
      less_than_1.5 = f(x %>% filter(distance_mi <= 1), y_var),
      less_than_1 = f(x %>% filter(distance_mi <= 0.5), y_var)
    )
  }
  
  data_df
}
```

Use **standard deviation (SD)** to measure how similar high schools are to each other.

- **Larger SD** = More spread / less similar
- **Smaller SD** = Less spread / more similar

# CA21-CA24

Compare the spread of all HS vs. HS close to border.

- SD of all HS (<b style="color: gray">gray</b> + <b style="color: red">colored</b> dots) = 21.04404
- SD of HS close to border (<b style="color: red">colored</b> dots) = 9.731259
- Delta = 21.04404 - 9.731259 = **11.31278**

**Positive delta** = HS closer to border are more similar

```{r}
load('../../cb_geomarket_shape/analysis_data/allyr_anal_eps_sf.RData')
ca_df <- hs_geomarket_distance_df %>% filter(border == 'CA21-CA24') %>% left_join(hs_df, by = 'hs_ncessch') %>% filter(!is.na(hs_pct_free_reduced_lunch))

leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(data = allyr_anal_eps_sf %>% filter(year == 2020, eps %in% c('CA21', 'CA24')), fillOpacity = 0, color = 'gray', label = ~paste0('<b>', eps, '</b>') %>% lapply(htmltools::HTML)) %>% 
  addPolylines(data = ca_df$geometry[[1]]) %>% 
  addPolygons(data = st_buffer(sf::st_set_geometry(hs_geomarket_distance_df %>% select(border, geometry) %>% filter(border == 'CA21-CA24') %>% distinct(), 'geometry') %>% st_transform(crs = 3857), dist = 2 * 1609.34) %>% st_transform(crs = 4326), color = 'gray', fillOpacity = 0.25, weight = 0, label = ~paste0('<b>', border, '</b>') %>% lapply(htmltools::HTML)) %>% 
  addCircleMarkers(data = sf::st_set_geometry(ca_df, 'hs_geometry'), radius = 2, color = if_else(ca_df$distance_mi <= 0.5, 'red', if_else(ca_df$distance_mi <= 1, 'darkorange', if_else(ca_df$distance_mi <= 2, 'gold', 'black'))), label = ~hs_sch_name)
```

```{r}
plot_eps('CA21-CA24', 'hs_pct_free_reduced_lunch')
```

```{r}
plot_lines <- function(data_df, title_text) {
  data_df %>%
    filter(!is.na(less_than_2)) %>% 
    mutate(delta = unrestricted - less_than_2) %>% 
    ggplot(aes(y = reorder(border, delta))) +
    geom_segment(aes(x = unrestricted, xend = less_than_2, yend = reorder(border, delta)),
                 color = "grey60", linewidth = 0.5) +
    geom_point(aes(x = unrestricted), color = 'grey30', size = 1) +
    geom_point(aes(x = less_than_2), color = 'red', size = 1) +
    geom_text(aes(x = (unrestricted + less_than_2) / 2,
                  label = sprintf("%+.2f", delta)),
              vjust = -0.2, size = 1.8, color = "grey30") +
    labs(title = title_text,
         subtitle = 'Gray = unrestricted, Red = less_than_2 | Labels show delta (unrestricted - less_than_2)',
         x = 'Standard Deviation', y = NULL) + 
    scale_y_discrete(expand = expansion(add = 1)) +
    theme_minimal() +
    theme(
      text = element_text(size = 10),
      plot.title    = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5, size = 10),
      axis.text.y = element_text(size = 6)
    )
}
```


# Percent free/reduced lunch

HS close to the border (<=2 mi) are equally or more similar to each other in:

214/(214+89) = **70.6%** of borders

```{r, fig.height=20}
lunch_data <- gen_agg_data('std', 'hs_pct_free_reduced_lunch')
lunch_data %>% filter(!is.na(less_than_2)) %>% mutate(delta = unrestricted - less_than_2) %>% mutate(x = delta >= 0) %>% pull(x) %>% table()

plot_lines(lunch_data, 'Standard deviation of all HS vs. HS within 2 mi of border')

ggsave(file.path(results_dir, 'comparison_lunch.png'))
```

# Zip median household income

HS close to the border (<=2 mi) are equally or more similar to each other in:

270/(270+59) = **82.1%** of borders

```{r, fig.height=20}
inc_data <- gen_agg_data('std', 'hs_zip_inc_house_med')
inc_data %>% filter(!is.na(less_than_2)) %>% mutate(delta = unrestricted - less_than_2) %>% mutate(x = delta >= 0) %>% pull(x) %>% table()

plot_lines(inc_data, 'Standard deviation of all HS vs. HS within 2 mi of border')

ggsave(file.path(results_dir, 'comparison_income.png'))
```
