---
#title: "The Road Not Taken: College Board Geomarkets and Recruiting Visits by Selective Colleges"
#title: "Recruiting as Performance: College Board Geomarkets and High School Visits by Selective Colleges"
#title: "Recruiting as Performativity: "
#title: "Performing College Recruiting: College Board Geomarkets and High School Visits by Selective Colleges"
#title: "The Performance of Recruiting: College Board Geomarkets and High School Visits by Selective Colleges"
#title: "Performativity in College Recruiting: College Board Geomarkets and High School Visits by Selective Colleges"
#title: "Performing Social Closure: College Board Geomarkets and Recruiting Visits by Selective Colleges"
#title: "Performing Social Closure: How College Board Geomarkets and the Market Segment Model Structure High School Recruiting Visits by Selective Colleges"
#title: "Performing Social Closure: How College Board Market Devices Structure High School Recruiting Visits by Selective Colleges"
title: "Performing Social Closure: How College Board Geomarkets Structure Recruiting Visits by Selective Colleges"
author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)", "Crystal Han (Stanford University)"]
bibliography: ./../../cb_geo/assets/bib/cb_geomarket_bib.bib
csl: ./../../cb_geo/assets/bib/chicago-author-date.csl
format:
  #docx:
  pdf:
    documentclass: article
    #papersize: letter
    number-sections: true
    number-depth: 2
    #toc: FALSE
    #pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #keep_tex: true
editor: visual
geometry:
  - margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(fixest)
#library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)

knitr::opts_chunk$set(echo = F, warning = F) 
options(knitr.kable.NA = '')

display_regression_table <- function(tbl_df, font_size = 10) {
  num_footer_rows <- 4
  
  tbl_df %>% 
    mutate(
      across(
        .cols = -term,
        .fns = ~ ifelse(
          term == "N",
          format(
            round(as.numeric(.), 0),
            big.mark    = ",",
            scientific  = FALSE,
            trim        = TRUE
          ),
          .
        )
      )
    ) %>% 
    kbl(
      booktabs  = TRUE,
      longtable = TRUE,
      # https://stackoverflow.com/a/49018919
      linesep = c(rep(c('', '\\addlinespace'), length.out = nrow(tbl_df) - num_footer_rows), rep('', num_footer_rows))
    ) %>% 
    kable_styling(
      font_size = font_size,
      latex_options = c('repeat_header', 'scale_down')
    ) %>%
    row_spec(nrow(tbl_df) - num_footer_rows, extra_latex_after = paste0("\\addlinespace\\cline{2-", length(tbl_df), "}")) %>% 
    footnote(
      general_title = '',
      general = 'Significance codes: *** p<0.001, ** p<0.01, * p<0.05',  # cutoffs specified in lpm.R
      threeparttable = TRUE,
      escape = FALSE
    )
}

#hey ho This many words: {{< words-total >}}
```

\pagenumbering{roman}
\newpage

**ABSTRACT**

Selective colleges are sites of social closure. Scholarship examines how selective colleges evaluate applicants, but colleges do not passively accept applications. They recruit. If college recruiting territories are structured by shared infrastructures, then mechanisms of social closure operate upstream of admissions. This article examines how College Board Geomarkets and the Market Segment Model structure high school recruiting visits by selective colleges. Geomarkets carve metropolitan areas into smaller geographic units meant to define local recruiting territories. The Market Segment Model predicts how student demand varies by Geomarket as a function of social class. These market devices became embedded in organizational structures and in third-party products that produce college recruiting behavior. Using data on recruiting visits made in 2017 by 42 selective colleges, we find that patterns of recruiting behavior are consistent with the Market Segment Model and that Geomarkets explain substantial variation in which high schools receive visits, even after controlling for school- and neighborhood-level covariates. The market devices that college admissions offices incorporate constitute a shared infrastructure that performs social closure – matching a socioeconomic hierarchy of communities to a socioeconomic hierarchy of colleges – prior to the evaluation of applications. We introduce the concept of infrastructural collusion: shared reliance by direct competitors on market devices that encode class hierarchy into the geography of opportunity.

<!-- ABSTRACT V1
Sociology of education investigates consumer-facing intermediaries (e.g., college rankings), but pays less attention to business-facing intermediaries that classify students on behalf of schools/colleges. The market devices literature offers useful concepts. Market devices are calculative tools incorporated into products or economic transactions. Performativity is the idea that actors utilize market devices in ways that enact behavior consistent with theories of market behavior. College Board Geomarkets and the Market Segment model are market devices for higher education recruiting. Geomarkets carve metropolitan areas into smaller geographic units meant to define local recruiting markets. The Market Segment model predicts how student demand for a college varies by Geomarket as a function of social class. These market devices were the basis for the College Board Enrollment Planning Service, an early software-as-a-service product that helped colleges plan recruiting. Using data on recruiting visits to high schools made in 2017 by a sample of 42 colleges, we examine whether the relationship between recruiting visits and Geomarkets is consistent with theoretical expectations about performativity.
-->

\newpage
\pagenumbering{gobble}
\pagenumbering{arabic}

# Introduction

<!-- V1 open: P1 = current lit on closure and raise a question; P2 = Geomarkets and how they are used 

Selective colleges are central sites of social closure. They confer access to scarce credentials and networks that facilitate entry into privileged occupations [@RN4935], while enrolling student bodies that remain sharply stratified by class and race [@RN4722; @abramitzky2024gi]. Weberian analyses emphasize the exclusionary strategies of status groups that redefine admissions criteria in ways that benefit the privileged group [@RN3544; @RN3547] and these admissions criteria diffuse across selective colleges through field-level mechanisms of mimetic and normative isomorphism [CITE BROADER LITERATURE]. Another literature traces legal and policy debates over race-conscious admissions [@warikoo_2025]. These explanations for social closure center the admissions process, but colleges do not passively accept applications; they recruit [@RN3519],  and the markets from which they recruit are not given — they are constructed. Therefore, a central question becomes how are recruiting markets delimited, commensurated, and rendered actionable. If college applicant pools are structured by shared recruiting infrastructures, then the mechanisms of social closure operate upstream of admissions.

College recruiting efforts depend on third-party "market devices" that classify high schools and communities on behalf of colleges looking for students. *The Structure of College Choice* [@RN4982] developed Geomarkets and the Market Segment Model. Geomarkes carve metropolitan areas into smaller geographic units meant to define local recruiting territories. The Market Segment Model categorized students into four market segments -- local, in-state, regional, and national -- based on the distance students were willing to travel to attend college. @RN4982 made the following observations: student market segments are a function of class; selective colleges primarily enroll students from the "regional" and "national" segments, who tend to have high-income, college-educated parents; the relative abundance of student market segments differs across Geomarkets; and the nature of competition between colleges socioeconomic homophily, whereby like-colleges compete for like-students on the basis of class. The Market Segment Model is the foundational logic of College Board's popular Enrollment Planning Service (EPS) software, which encouraged colleges to identify recruiting territories by investigating which Geomarkets demonstrated strong student demand for peer colleges. Further, Geomarkets became an organizing principle for how colleges allocate admissions personnel to recruiting territories [CITE]. We argue that Geomarkets and the Market Segment Model facilitate social closure by creating an infrastructure that encourages selective colleges to concentrate their recruitment efforts within affluent Geomarkets. This study examines how Geomarkets and the Market Segment Model structure high school recruiting visits by a sample of 42 selective colleges. 
 -->
 
> The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one’s peers [@RN4982]

Selective colleges are sites of social closure. They confer access to scarce credentials and networks that facilitate entry into privileged occupations [@RN4935], while enrolling student bodies that remain sharply stratified by class and race [@RN4722; @abramitzky2024gi]. @RN4982 observes that competition between colleges occurs within "collective markets," defined by socioeconomic homophily, whereby like-colleges compete for like students on the basis of class. The Consortium on Financing Higher Education (COFHE) -- a group of selective, private colleges -- is one such collective market. "What every envious admissions officer outside the Consortium understands," @RN4982[p. 45] write, "is that these thirty institutions are strengthened by their very collectivity, though they are in fact one another's principal competitors." With support from College Board, Zemsky developed these observations into market devices for colleges that intervene on the college choice process of students. 

Existing explanations of social closure emphasize exclusionary strategies of status groups that redefine admissions criteria in ways that benefit the privileged group [@RN3544; @RN3547] and then diffuse to other colleges [CITE BROADER LITERATURE]. Another literature traces legal and policy debates over race-conscious admissions [@warikoo_2025]. These explanations center the admissions process. But colleges do not passively accept applications. They recruit [@RN3519]. The markets they recruit from are not given, they are constructed. If college recruiting territories are structured by shared recruiting infrastructures, then the mechanisms of social closure operate upstream of admissions, raising questions about how recruiting territories are defined and made actionable for colleges.

College recruiting efforts depend on third-party "market devices" that classify high schools and communities on behalf of colleges looking for students. Market devices are tools that calculate and classify. They become incorporated into organizational goals and routines in ways that produce a menu of possible and likely actions. Market devices engender classification situations @RN4810], whereby a vertical hierarchy of products is matched to a vertical hierarchy of consumers. We argue that Geomarkets and the Market Segment Model -- developed by Zemsky with support from College Board -- are market devices that match "collective markets" of selective colleges to affluent communities, performing social closure upstream of admissions. 

Geomarkes carve metropolitan areas into smaller geographic units meant to define local recruiting territories. The Market Segment Model categorized students into four market segments that are a function of class. @RN4982 argue that selective colleges primarily enroll students from high-income, college-educated households, but the relative abundance of these students differs across Geomarkets. Over time, Geomarkets and the Market Segment Model became embedded in college organizational structures and in shared infrastructure colleges draw from. Because affluent Geomarkets concentrate the students selective colleges are looking for, this infrastructure systematically directs recruiting attention toward wealthy communities.

This study examines how Geomarkets and the Market Segment Model structure high school recruiting visits by a sample of 42 colleges -- 14 selective private research universities, 12 selective private liberal arts colleges, and 16 public research universities. While selective private colleges are understood as instruments of intergenerational class transmission, public research universities have been engines of social mobility. However, state disenvestment encourages public research universities to search for nonresident tuition revenue in the same set of affluent Geomarkets as their private peers. 

Analyses are guided by three research questions. First, to what extent are visit patterns consistent with recommendations from the Market Segment Model? To what extent do Geomarkets explain which high schools receive recruiting visits? How does Geomarket popularity -- measured as number of visits per high school -- mediate the relationship between school socioeconomic and racial/ethnic composition and the probability of a school receiving a visit?

For the selective private -- and most public -- colleges in our sample, the Market Segment Model recommends focusing on affluent Geomarkets located in regions where students are likely to consider a college like yours [@RN4982]. Although a small number of (mostly public) colleges follow an in-state recruiting strategy, most colleges in our sample follow a national recruiting strategy, in which the majority of visits are to schools outside their home region (e.g., New England). Visits by private colleges -- and out-of-state visits by public universities -- are highly concentrated in affluent Geomarkets. With respect to RQ2, fixed effects regression models find that Geomarkets explain substantial variation in recruiting visits even after controlling for a strong set of school- and neighbohood-level controls. With respect to RQ3, regression models find that schools with a low share of free/reduced lunch students benefit more from being located in a popular geomarket than schools with a high share of free/reduced lunch and schools with a modest share of Black/Hispanic students benefit more from being located in a popular Geomarket than schools with a high share of Black/Hispanic enrollment. Drawing from @RN4808, a wealthy, preominantly white school in a wealthy, predominantly white Geomarket is a "good match" with what selective colleges expect.<!--, whereas poor, predominantly Black/Hispanic school in a wealthy Geomarket is not a good match -->

We make targeted contributions to the sociology of education and a broader contribution to the mechanisms of social closure. Scholarship on selective college access focuses on conflict over admissions criteria and who benefits [CITE], while scholarship on enrollment management shows that colleges devote substantial resources identifying and recruiting applicants [CITE]. Both literatures center colleges as protagonists, ignoring business-facing intermediaries that classify students on behalf of colleges. Our research points to an upstream ecosystem of third-party vendors, products, and professional associations that that collectively construct the infrastructure that colleges utilize to plan and implement recruiting strategy. Geomarkets and the Market Segment Model were grafted atop the history of racial and class segregation and became embedded within third-party products and college organizational structures. These market devices contribute to social closure by facilitating a tiered matching of colleges to places, encouraging selective private colleges to target high-income localities nationwide, while working class communities are left to less selective institutions. Although public research universities were engines of social mobility in the era of strong state funding, the desire to increase nonresident enrollment encourages to encourages public research universities to target the same set of affluent Geomarkets coveted by selective private institutions. <!-- delete cuz redundancy. Therefore, Geomarkets and the Market Segment Model contribute to social closure by engendering an infrastructure that matches schools and communities to colleges based on socioeconomic privilege. -->



The manuscript is organized as follows. First, we review scholarship on college access and enrollment management. Second, we describe the case of Geomarkets and the Market Segment Model. Third, we develop a theoretical framework based on market devices and classification situations. Fourth, we describe data and methods. Fifth, we present results. Finally, we discuss implications for scholarship. Scholarship at the intersection of quantification and education has considered consumer-facing intermediaries (e.g., college rankings) that discipline consumers and, in turn, businesses [CITE]. However, intermediaries that classify consumers on behalf of businesses are more pervasive and more sophisticated [@RN5015]. In the education industry, these intemediaries match ordinally arranged consumers to ordinally arranged products. We argue that sociology of education can develop novel insights by shifting the object of inquiry from students and direct providers education (schools, and colleges) to the market devices that classify students for direct providers.

# Literature Review: College Access in the Enrollment Economy

Social closure, associated with Max Weber [CITE], is the process by which advantaged groups maintain their privilege by restricting access to resources and opportunities that confer economic and social privilege. Weber conceived of social closure as competition between status groups, defined of "associational communities comprised of all who share a sense of status equality" [@RN3547, p. 3] based on membership of a characteristic (e.g., racial, ethnic, religious, social) that may be the basis of exclusion. Following the transition to rational-legal authority defined by universalistic bureacracies, organizations that confer economic and social privileges become the site of contestation between dominant groups seeking exclusionary closure and marginalized groups seeking usurptionary closure. Colleges are the exemplar site of contestation [@RN865]. Weber states, 

> The development of the diploma from universities \ldots make for the formation of a privileged stratum in bureaus and in offices. Such certificates support their holders’\ldots claims to monopolize socially and economically advantageous positions. When we hear from all sides the demand for an introduction of regular curricula and special examinations, the reason behind it is, of course, not a suddenly awakened ‘thirst for education’ but the desire for restricting the supply for these positions and their monopolization by the owners of educational certificates\ldots As the acquisition of the educational certificate requires considerable expense and a period of waiting for full remuneration, this striving means a setback for talent (charisma) in favor of property.

In the U.S., selective colleges confer access to privileged social and economic positions. As college access expands to previously excluded populations, scholarship in sociology and economics document growth in the benefits of graduating from a selective institution [@RN4935; @RN2190], and growth in competition for access to selective institutions [CITE]. Based on data from the IRS, the "Mobility Report Cards" [@chetty2020mobility] measures both parental household income and student earnings post-graduation at the college level for students who were born in 1991 and were freshman around 2009. Although low-income students students who graduate from selective colleges have strong labor market outcomes [@RN4722], access to selective institutions is dominated by weealthy households. For example, at Colorado College, one of the colleges in our analysis sample, median family income was about $379,000 (2026 CPI), with just 2% of freshman coming from the bottom income quintile, 78% of freshman coming from the top quintile, 54% of freshman from the top 5%, and 24% of freshman from the top 1% [@chetty2020mobility].

Several literatures in the sociology of education have analyzed the admissions criteria utilized by selective colleges and which applicants these criteria benefit. The Weberian status group social closure tradition is articulated by the historical sociology of [@RN3544; @RN3547]. Leaders of Ivy League institutions were themselves members of the elite Protestant status group and had an interest in blocking Jewish enrollment growth by developing "character" admissions criteria associated with distinctly upper-class Protestant activities and traits. [TRANSITION?] A robust literature analyzes the rise and fall of affirmative action [@RN4874; @RN2294; @warikoo_2025; @RN3541]. Another literature ties the growing importance of SAT/ACT scores in selective college admissions to the rise of college rankings systems that reward colleges for enrolling high-scoring students [@grodsky_warren_felts_2008; @RN1749; @RN5003]. Recent scholarship examines the implementation of "test optional" policies [@RN5103] and holistic admissions processes that evaluate applicants with respect to the opportunities available in their local context [CITE BASTEDO]. Although these studies offer rich accounts of how colleges evaluate applicants, the scholarly preoccupation with selective admissions obscures the basic fact that overwhelming majority of colleges -- including most selective colleges -- cannot thrive by passively receiving applications.

<!-- PREVIOUS TRANSITION SENTENCE 
Although the selective admissions literature offers rich accounts of how colleges evaluate applicants, it offers less insight into how colleges identify and cultivate prospective applicants before admissions evaluation occurs.
-->

The "enrollment economy" refers to reliance on student enrollment for organizational stability. The concept was introduced by @RN2401 \[p. 332\], in his analysis of California adult education institutions: "The most important pressures on these schools in their day-to-day administration arise from the *enrollment economy* \ldots Financial support from the state is figured by the hours of attendance logged the previous year \ldots and a major slump in attendance constitutes a serious threat to organizational welfare." (p. 332). At the opposite end of the prestige spectrum, @RN1549 [p. 17] describes dependence on students by selective private colleges: "the technology of producing much of what is sold in higher education is unusual in that colleges can buy important inputs to their production only from the customers who buy their products; that is, higher education uses a customer-input technology. Colleges depend on enrolled students for tuition revenue and on alumni for donation revenue. They depend on enrolled students for peer learning effects. Furthermore, the characteristics of enrolled students -- for example, average SAT scores -- substantially determine their national ranking, which then determines student demand for the subsequent cohort of prospects \[CITE\] and also affects research funding and donation revenue \[CITE\].

Organizational stability in the enrollment economy is the responsibility of enrollment management (EM). As a profession, EM integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]. As an administrative structure, the "office of enrollment management" typically controls and coordinate the functional activities of admissions, financial aid, and recruiting [@RN2406; @RN3625].^[@RN2406 exemplifies a literature at the intersection of EM and organizational behavior that shows how concerns about the enrollment economy motivate changes in organizational behavior, such as diversifying degree programs or changing the name of the organization [@RN869; @RN515; @RN2518]] The administrative structure of EM offers one means of categorizing scholarship about the supply-side of college access. Accross disciplines, the majority of scholarship focuses on admissions [e.g., @RN4874; @RN4993; @RN4131; @RN3522] or on financial aid [e.g., @RN3133; @RN3762]. These literatures offers less insight into how colleges identify and cultivate prospective applicants before admissions evaluation occurs. Therefore, a growing literature in sociology analyzes the earlier "recruiting" stages of acquiring "leads," and soliciting "inquiries" and applications \[CITE\]. Several monographs analyze recruiting as part of broader studies on enrollment management [@RN3519], high school to college transitions [@RN4324], elite reproduction [@RN4407], or political economy [@cottom2017lower]. From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options. @cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.

@RN3519 provides an ethnography of enrollment management at a selective private liberal arts college. Being tuition-dependent and sensitive about its position in the rankings, The College focuses on enrollment goals related to academic profile and revenue generation. @RN3519 highlights the relational function of off-campus recruiting visits: "the College's reputation and the quality of its applicant pool are dependent upon its connections with high schools nationwide" (p. 54). During the autumn "travel season," admissions officers visit selected high schools across the country "to spread word of the institution and maintain relationships with guidance counselors" (p. 53-54). @RN4741 reported that "calling high school counselors after visits" was an effective and widely-used outreach strategy for both private universities and public universities. Thus, relationship maintenance remains a high-touch process even in the digital age. The College tended to visit the same "feeder" schools year after year because recruiting depends on long-term relationships. Visited schools tend to be affluent schools -- in particular, private schools -- that possess the resources to host a visit and enroll high-achieving students who can afford tuition.

Whereas scholars in the sociology of education have investigated recruiting as part of a broader analysis, AUTHORS [-@RN4758; -@RN4759; -@RN5023] have attempted to develop a literature that makes college recruiting behavior the object of empirical analysis. @RN4758 argue that off-campus recruiting visits are a credible indicator of college enrollment priorities.[^2] [INSERT SENTENCES SAYING HOW MUCH IS SPENT ON OFF-CAMPUS RECRUITING AND SCHOLARSHIP ON THE EFFECTS OF OFF-CAMPUS RECRUITING]. @RN4758 analyzed recruiting visits made by public research universities. The authors found that 12 of the 15 universities made more visits to out-of-state schools than in-state schools. These out-of-state visits focused on affluent, predominantly white public high schools. Set against scholarship about nonresident enrollment growth as a response to declining state support [@RN3753; @RN3685], these results suggest that many public research universities prioritize enrollment from wealthy out-of-state students.

[^2]: Sociological theories of organizational behavior argue that organizations expend substantial resources on goals they care about while symbolically adopting -- highly visible, ceremonial behavior with few resources expending -- goals demanded by stakeholders in the broader institutional environment [CITE].

@RN5023 conduct a social network analysis of recruiting visits to private high schools by a sample of public research universities and selective private research universities. A visit indicates a social relationship between a college and a high school, suggesting that the college wants to enroll some students from the high school and the high school wants to send some students to the college. Both public and private universities made a disproportionately large number of visits to private schools relative to public schools. Analyses revealed high overlap in the recruiting networks of public and private universities. @RN4759 conducted geospatial analyses of out-of-state recruiting visits to Los Angeles and Dallas by four public research universities. She found that universities engage in "recruitment redlining -- the circuitous avoidance of predominantly Black and Latinx communities along recruiting visit paths" (p. 585).

<!-- PARAGRAPH I CUT CUZ IT SEEMED LIKE A DISCLAIMER THAT KILLED THE ENERGY

Upon review, prior scholarship on the supply-side of college access identifies multiple mechanisms that contribute to social closure in who attends which colleges. Although colleges do not respond uniformly to shared pressures [e.g., @RN5103], the literature also documents broad similarities in the enrollment management behavior of peer colleges and colleges from different populations [e.g., @RN2406; @RN5023]. Scholarship on organizational behavior argues that organizations within a field behave similarly because they face similar external pressures and institutional environments [@RN527; @bastedo2011rankings]. For example, prior research describes college rankings as an emergent field-level institution that disciplines prospective students’ search behavior and employers’ hiring behavior and, in turn, shapes colleges’ enrollment management strategies.
-->

Upon review, scholarship locates inequality in college access in the behavior of households, schools, and colleges [CITE]. Scholarship on the supply-side of college access -- selective admissions, EM, recruiting -- argues that students are sorted by the actions of colleges, which are conditioned by macro-institutions in the organizational field [CITE]. This focus on college behavior understates the role of third-parties that mediate between consumers and sellers. Sociology of education has investigated intermediaries that "occupy the interface between consumers and producers" [@RN5065, p. 1], including scholarship on college rankings [@RN5003] and school rankings [@RN5005; @RN5075; @hasankumar2019Digitization]. To our knowledge, intermediaries that classify consumers on behalf of sellers have rarely been the object of empirical research [but see @RN5039]. Although consumer-facing intermediaries are salient to media outlets and researchers, intermediaries that classify consumers for businesses are more sophisticated, pervasive [@RN4957; @RN5064; @RN5073; @RN4808], and may be highly salient to professionals responsible for recruiting customers. Scholarship on college access has ignored the infrastructural mechanism of third-party enrollment management tools that classify students and localities on behalf of colleges. These tools -- which are built atop the foundation of segregation in American communities -- facilitate a tiered matching of colleges to places by encouraging selective private colleges and public research universities target high-income localities nationwide, while encouraging less selective institutions to target nearby working- and middle-class communities.

Geomarkets were conceived in service of a class-based model of college choice [@RN4982]. Subsequently, Geomarkets became the basis of College Board Enrollment Planning Service (EPS) software that helped colleges plan recruiting efforts and an organizing principle allocating admissions personnel to recruiting territories. We conceptualize these tools as "market devices" [@callon1998embeddedness]  that create "classification situations" [@RN4810], whereby an ordinal hierarchy of students is matched to an ordinal hierarchy of postsecondary institutions. Performativity is the idea that market devices are used in ways that create the behaviors they predict [@RN5067]. Empirically, we show that Geomarkets explain substantial variation of high school visits by a sample of selective private colleges and public research universities, even after controlling for a strong set of school- and neighborhood-level covariates. Finally, we contribute to a critical sociology of the College Board. Existing scholarship focuses on College Board's consumer-facing testing products \[CITE\], mirroring the blind-spot observed in scholarship on intermediaries. This manuscript shows that College Board testing products are the basis for business-facing products that classify prospective students on behalf of colleges.


```{=html}
<!-- CUT TEXT AND OG PARAGRAPHS

To our knowledge, this is the first empirical analysis in the sociology of education based on concepts from the market devices literature. These concepts offer great utility for future research that investigates the growing role of third-party vendors in schools and colleges. 

- ozan says; seems like I can make the "first to use market devices" contribution elsewhere. better in the discussion or intro I think.
- GPT says: Yes — I agree. Put “first to use market devices” in the intro (end of intro / contribution paragraph) or the discussion, not here.

Finally, analyses for RQ3 make a targeted contribution to the classification situations literature [@RN5015] by assessing idea of "good matches" between producers and consumers.

Scholarship about third-parties that mediate between consumers and producers/sellers considers two broad types of intermidiaries: intermediaries "that occupy the interface between consumers and producers" [@RN5065, p. 1], including college rankings and expert critics; and intermediaries akin to the "classification situations" of @RN4810 that classify consumers on behalf of sellers. We might cede that consumer-facing product ratings have more pathways of influence; they discipline consumer preferences, which in turn disciplines product attributes created by producers. On the other hand, intermediaries that classify consumers for the sake of producers -- including the entire field of market research -- are much more sophisticated [@RN4957; @RN5064; @RN5073; @RN4808]. To our knowledge, scholarship on quantification in the sociology of education has mostly focused on consumer-facing intermediaries. Certainly college rankings [@RN5003] and school rankings [@RN5005; @RN5075; @hasankumar2019Digitization] are salient for decision-makers and have important consequences. Perhaps another reason for the volume of scholarship about consumer-facing intermediaries is that coverage by the popular press makes them known to researchers. Intermediaries that classify consumers for businesses receive scant attention from the popular press but may highly salient for professionals who are responsible for identifying and recruiting customers. 

Robust literatures in the sociolology of organizations [@RN4421] and the sociology of education [@fast2025racializeddecoupling; @RN5055; @RN4873] shows how the behavior of organizations reproduces racial, class, and gender inequality \[CITE\]. @RN5070 contributes to scholarship on gender inequality because scholarship that locates inequality in organizational behavior fails to observe gender inequality produced by the market. @RN4873 casts colleges as sorting machines that incorporate and reproduce social inequality. This line of research fails to consider how the market devices utilized by colleges incorporate and reproduce social inequality. Whereas @RN5070 shows that market devices produce social inequality by directly classifying consumers, this manuscript investigates how market devices reproduce inequatlity of opportunity by disciplining how colleges classify consumers.

-->
```


# The Case: Geomarkets and the Market Segment Model

```{=html}
<!-- 
We describe the case of Geomarkets and the Market Segment Model, adding deeper context as we introduce concepts from market devices, commensuration, and ordinal classification. -->
```

Geomarkets were conceived in the late 1970s, when colleges were concerned about declining demographics. @RN4982 \[p. ix\] write that demographer "Stephen @RN5076 had already frightened everyone who would listen with his projection of 40-percent declines in college enrollments by the close of the 1980s." Elite colleges responded by leaning on the associational configuration [@RN4517] to solve problems common to the group. The project began when UPenn president, Martin Meyerson, tasked UPenn professor Robert Zemsky, to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?'" [@RN4982, p. x]:

> "In pursuit of these questions, we began working in 1978 with the Market Research Committee of the Consortium on Financing Financing Higher Education (COFHE). The Consortium, a group of thirty selective private institutions, had in its brief history developed a remarkable tradition of sharing data\ldots For nearly a year, the Consortium's Market Research Committee provided focus to our efforts. Much of this project's framework -- the Market Segment Model and its derivation from admissions officers' folklore was shaped and then reshaped in our discussions with members of the committee."

```{=html}
<!-- FULL QUOTE
> "In pursuit of these questions, we began working in 1978 with the Market Research Committee of the Consortium on Financing Financing Higher Education (COFHE). The Consortium, a group of thirty selective private institutions, had in its brief history developed a remarkable tradition of sharing data. Members of COFHE had come to know almost as much about one another's applicant pools as they did about their own. For nearly a year, the Consortium's Market Research Committee provided focus to our efforts. Much of this project's framework -- the Market Segment Model and its derivation from admissions officers' folklore was shaped and then reshaped in our discussions with members of the committee."
-->
```

@RN4982 conceived of the Market Segment Model as an effort to "capture and quantify" (p. 11) the knowledge of admissions officers: "We believe that the intuitions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research is based upon listening carefully to what admissions officers have to say (pp. 9-10). The "initial \[analytic\] task" was to define geographic boundaries "in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982, p. 4]. Quoting an admissions officer, @RN4982 \[p. 11\] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982 [p. 11] created "three types of boundaries -- region, state, and community." The initial regions in the project were New England, Middle States, and the South. Next, they "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p. 11). These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors. @RN4982 [pp. 11-12] described the creation of Geomarket borders only briefly:

> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication. Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs. In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful.

Over time, the project goal morphed into predicting student demand for any college from any locality. However, "to gain a truly comprehensive view of the collegiate enrollment market, we needed a database that described most institutions and most students" [@RN4982, p. x]. The group approached College Board. "Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p. x]. College Board provided score-sending data for all 1980 SAT test-takers. For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region). @RN4982 classified each SAT test-taker into one of four student *market segments*. A student is categorized in the "local" market segment if they submit more SAT scores to local institutions than they do to in-state, regional, or national institutions. An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc. The Market Segment Model "was nothing more than a set of simple rules for disaggregating high school seniors into similar groups" [@RN4982, p. 4]. "The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways."

@RN4982 [Chapter 3] investigates which student characteristics predict whether a student belongs to the local, in-state, regional, or national market segment. The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score-sending behavior, both individually and in combination. The resulting thesis is that student demand for college is a function of social origin. These four variables "reflect the basic social patterns of the nation; it would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p. 33]. The analyses conclude (p. 42) that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations."

However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics. As college leaders sought new sources of patronage, the Market Segment Model yielded practical implications about which communities to target. @RN4982 [p. 44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:

> On occasion, senior spokespersons for the profession worry that students outside the main \[Geo\]market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.

@RN4982 developed two analytic outputs, (1) the Market Segment Profile and (2) the Institutional Profile. The Market Segment profile showed -- separately by Geomarket -- the number of students from each market segment (local, in-state, regional, national), their socioeconomic characteristics, and average test scores. The Institutional Profile showed -- separately by Geomarket and institution (your own or a competitor) -- how many students from each market segment sent scores to that institution.

In 1984, College Board transformed the Market Segment Model and associated outputs into the Enrollment Planning Service (EPS), an early software-as-service platform sold to colleges that was based on @RN4982. For any Geomarket, colleges could obtain the Market Segment Report and the Institutional Profile for themselves or a competitor. Based on background conversations with EM professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket. Typical @RN5019 marketing material describes EPS as, "the marketing software that pinpoints the schools and Geomarkets where your best prospects are most likely to be found. With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition." Following @RN4982, the EPS platform encouraged users to, first, identify promising Geomarkets and, second, to identify desirable high schools for recruiting visits. To find new markets, EPS encouraged colleges to select peer colleges and identify Geomarkets where a large number of students sent scores to those colleges. Market research suggests that EPS was highly salient. @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.


# Theory: Market Devices and Classification Situations

<!-- OG TRANSITION SENTENCE *Market devices*. We conceive of Geomarkets, the Market Segment Model, and the subsequent Enrollment Planning Service product as examples of "market devices" [@RN5064; @RN5067]. 
-->
The market devices literature examines how the components of markets are assembled and enacted, and how these processes shape behavior [@callon1998embeddedness; @RN5064; @RN5067]. Market devices are discreet calculative tools (e.g., algorithms, metrics, graphs, rankings models, spreadsheets) that are incorporated into products or economic transactions. @RN5064[p. 2] define market devices as a "way of referring to the material and discursive assemblages that intervene in the construction of markets." Here, material refers to tangible tools and technologies, like spreadsheets, algorithms or software, while discursive refers to representational elements, such as categorical, quantitative, and ordinal variables. Assemblages refers to a bundle of heterogeneous components that work together, suggesting that market devices consist of and build upon other market devices. Geomarkets are a market device -- based on the SAT -- that helps enrollment managers define recruiting territories and make decisions about which high schools to recruit from. Market devices are non-neutral, actively shaping outcomes. Therefore, markets do not exist a priori waiting to be observed; they are constructed by market devices. From the market devices perspective, the transformation of American higher education "from a collection of local autarkies to a nationally integrated market" [@RN2133, p. 1] can be (partially) explained as the consequence of enrollment managers using Geomarkets to achieve enrollment goals.

Theories (e.g., pricing theory, status attainment theory), by themselves, are not market devices. They are abstract ideas that can be applied to specific contexts. For example, homophily is the idea that people who share similar characteristics are more likely to connect [@mcpherson2001birds]. Theories become part of market devices when components of the theory are embedded within the device. @RN4982 argue that homophily is the organizing principle of competition between colleges; similar colleges compete with one another for similar students. An analytic output from @RN4982 is the Institutional Profile, which provided information about students from a particular Geomarket who sent SAT scores to a particular institution. This functionality was subsequently embedded in EPS software, which encouraged colleges to investigate which Geomarkets have high demand for peer colleges

In the market devices framework, neither market devices nor actors have agency in isolation from one another. Whereas market devices are tools, "agencements" are "socio-technical configurations" [@RN5102] of market devices, people, and organizational arrangements (goals, roles, routines, incentives) designed to produce a menu of possible actions and to make some actions more likely than others. Agencement orients research to the intersection of what the market devices is capable of doing and the organizational context that is utilizing the device -- in concert with other devices -- to achieve organizational goals. Geomarkets are an input for enrollment management offices charged with realizing college enrollment goals. Geomarkets can be utilized to define recruiting territories and assign admissions officers to territories. EPS can be used to investigate which Geomarkets exhibit strong demand for the college, for peer institutions, and to plan which high schools in a Geomarket will be visited by which admissions officer.
<!-- To make these concepts concrete, consider the case of US News College Rankings. The market device is the ranking tool itself -- which consists of data inputs, an algorithm, and a ranked list published on a website. Agencements refer to the ways that actors utilize the rankings market device. For example, prospective students incorporate rankings in their college search process, employers use rankings to decide which colleges they will recruit from, and college administrators make resource allocation decisions designed to move up in the rankings [@RN5003].
-->

Performativity is the idea that theories of market behavior do not merely describe behavior, rather market devices are constructed and acted upon in ways the bring about behavior consistent with the theory [@RN5064; @RN5067]. For example, @RN5067 provide a historical sociology of the Chicago Board Options Exchange, which opened in 1973. An option is a contract that gives the holder the right to buy or sell an asset (e.g., a stock) at a given price. Economists Fischer Black and Myron Scholes developed the Black-Scholes model -- a calculative market device -- to predict the price of options @RN5074, eventually winning the Nobel Prize in Economics. Initially, the Black-Scholes algorithm did not accurately predict the price of options. Black-Scholes model predictions became more accurate over time because options traders began embedding the logic and components of the model into spreadsheets/software. That is, the models predictions became more accurate because the model produced the prices it predicted [@RN5067]! @RN1776 provide an example of performativity for the case of US News College Rankings. The reputational assessment of senior administrators was a substantial component of the rankings algorithm. @RN1776 show that administrators based their assessment scores on the prior overall rankings. Therefore, the rankings algorithm disciplines administrators to provide assessments that are consistent with the prior rankings.

Although canonical contributions to the market devices literature did not focus on structural inequality -- race, gender, class, @RN4777 argue that market devices literature can offer novel insights about the mechanisms of structural inequality. Analysis of the Houston residential real estate market by Korver-Glenn [-@RN4801;-@RN4779] shows how market devices can be a mechanism of racial inequality. Every home listing is a assigned to a Houston Association of Realtors (HAR) "market area" -- a market device akin to Geomarkets -- based on the address of the property. In 70 of the 98 HAR market areas, one racial group comprised at least 50% of residents. @RN4779[p. 641] observed that "50 of these 70 market areas were majority-white -- a disproportionately high number given that only 31 percent of the population within the encompassing area is white." On average, majority-white market areas contained about 20,000 residents, while majority-Black and majority-Hispanic market areas, respectively, contained 61,000 and 88,000 residents. Majority-white market areas also exhibited greater socioeconomic homogeneity. Because appraisals are based on recent comparable sales within the same market area (agencement), homes in high-income white neighborhoods received higher appraisal values than homes in high-income Black or Hispanic neighborhoods. College Board Geomarkets often exhibit similar differences in granularity. For example, the affluent, predominantly White "IL 9 - North Shore" Geomarket, had a 2020 population of about 161,000 while "IL 12 - South and Southwest Suburbs" Geomarket had a 2020 population of about 1.2 million.


<!-- CUT TEXT SUMMARIZING KRIPPNER "GENDERED MARKET DEVICES" AS AN EXAMPLE OF MARKET DEVICES AND STRUCTURAL INEQUALITY; KORVER-GLENN IS A MORE APPROPRIATE EXAMPLE FOR GEOMARKETS; KRIPPNER IS MORE APPROPRIATE FOR GEODEMOGRAPHIC SEGMENT CUZ BOTH KRIPPNER AND SEGMENT ARE BASED ON CLUSTER-TYPE ANALYSES

@RN5070 offers an empirical illustration in her analysis of insurance products. @RN5070 [p. 598] observes that prior feminist organizational scholarship considers gender discrimination an "organizational matter -- a product of administrative routines such as hiring procedures, promotion protocols \ldots \[but\] this approach reinforced the notion that markets do not themselves contribute to discriminatory outcomes." Drawing on documents from lawsuits filed by the National Association for Women, @RN5070 provides a historical document analysis that centers on the statistical models used to create life and car insurance products. She describes these models as "gendered market devices" because gender is one of the input variables utilized to assign individuals to different "risk pools."
-->


The concept commensuration [@RN5071; @RN5063] helps us gain deeper insight about the agencement of market devices created by College Board. Commensuration is "the expression or measurement -- of characteristics normally represented by different units -- according to a common metric" [@RN5063, p. 315]. Commensuration is a process that requires work, akin to institutionalization. Once achieved, commensuration reduces the cognitive effort of decision-making. It is "a way to reduce and simplify disparate information into numbers that can easily be compared. This transformation allows people to quickly grasp, represent, and compare differences" (p. 316). For example, college rankings enable prospective students to compare colleges based on a single metric. Therefore, commensuration "can be understood as a system for discarding information and organizing what remains into new forms" [@RN5063, p. 317]. @RN5005 analysis of UK school "league tables" -- a commensurating market device that facilitated comparing schools on the basis of test scores -- shows how commensuration can be a mechanism of social reproduction; professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools. <!-- detail on McArthur and Reeves
^[In 1992, the UK government began publishing school league tables that ranked schools on student performance on national exams taken at age 16. League tables disciplined households -- and in turn schools -- to conceive of school quality in terms of test scores. League tables facilitated making evaluative comparisons between schools regardless of geographic proximity, but only in terms of the test score metric that was substantially a function of the class composition of schools. Following the adoption of league tables, localities with higher performing schools experienced growth in the share of managerial/professional households and a decline in the share of working-class households [@RN5005].]
-->

<!-- CUT PARAGRAPH ON SAT AS A COMMENSURATING MARKET DEVICE
Economics conceives of the SAT as a market device that facilitates commensuration. @RN2247 argues that "information costs" were a barrier to the emergence of a national higher education market that efficiently matches students to colleges: "In 1955, there was no early national college aptitude test. Students and colleges simply did not know where students stood in the national distribution of high school graduates’ achievement or aptitude" (p. 103). Colleges could not make an apples-to-apples comparison between students from different schools because the information on a high school transcript "is relative to a standard that a college will not understand unless it draws very often from the high school" (p. 103). @RN2247 credits the SAT/ACT exams for the emergence of an efficient, national higher education market. Institutionalization of the SAT/ACT caused a "dramatic fall" in the cost of "colleges’ information about students" [@RN2247, p. 102].
-->


College Board Geomarkets are a commensurating market device that was built on the SAT and became an input into subsequent market devices. Drawing from @RN2247, the SAT is a commensurating market device that enabled colleges to compare students from different schools, causing a "dramatic fall" in the cost of "colleges’ information about students" (p. 102). The Market Segment Model is a theory of student demand that categorizes students into four market segments -- local, in-state, regional, and national -- based on SAT score-sending behavior. Geomarkets are a market device that defines geographic localities and enables colleges to commensurate these localities based on the number of students from each market segment. Whereas UK league tables allowed parents to compare geographically disparate schools based on test scores [@RN5005], Geomarkets and the Market Segment Model enables enables colleges to compare geographically proximate and geographically disparate Geomarkets based on the number of students from each market segment. Each student market segment is associated with particular sets of institutions. @RN4982 states that selective colleges primarily draw students from the regional and national market segments. The key to finding new student markets, then, is identifying Geomarkets with large numbers of regional and national students. Geomarkets are then incorporated into the Enrollment Planning Service product, enabling colleges to decide which Geomarkets to recruit from and which high schools to visit [@RN4980].

**Classification situations.** Scholarship on classification situations [@RN4810; @RN5015; @RN4808; @RN5072] helps conceptualize how Geomarkets faciliate social and racial stratification in college access. Traditionally, sociology defined social class from the production side of the economy (e.g., worker vs owner; clerical, manager, professional). By contrast, @RN4810 argue that processes of ordinalization "endogenous to the market" (p. 560) define class through effects on the *consumption* side of the economy.  

Classification situations are ordinal interventions whereby organizations "sort and slot people into new types of categories with different economic rewards or punishments attached to them" [@RN4810, p. 561]. The canonical example is credit scores [@RN4877], which enable third-party intermediaries on supply side to "slice consumers into behaviorally-defined risk groups, and price offerings to them accordingly. On the demand side, consumers find themselves \ldots fitting into these categories -- which, by design, are not constructed from standard demographic classifications such as race" [@RN4810, p. 560] but are highly correlated with social categories. The goal of classification situations is "good matches," whereby an ordinal hierarchy of products match to an ordinal hierarchy of people. Businesses "seek good matches when they want to link a wealthy customer with a quality credit card, or a heavy drinker with a bad insurance plan" [@RN4808, p. 17]. Whereas economics tends to view such matches favorably -- for example, @RN2247 [p. 106] writes the "efficient outcome allocates students to colleges based on students’ ability to benefit from the \ldots magnitude of the human capital investment that the college offers" -- Sociology tends to highlight the predatory, extractive inclusion of formerly excluded consumers (e.g., for-profit colleges) [@RN4774].


<!-- CUT SENTENCE
@RN5072 argues that the transformation from nominal to ordinal categories serves political liberalism because ordinal classifications are more inclusive. 
-->

```{=html}
<!-- PUT THIS BELOW INSTEAD

> Companies seek good matches when they want to link a wealthy customer with a quality credit card, or a heavy drinker with a bad insurance plan. This kind of matching capitalizes on people’s differentiated dispositions and practical habits, what Pierre Bourdieu calls their habitus. The ‘good matches’ are perceived as natural because they fit well with how things already are.
-->
```

<!-- The Market Segment Model is a class- and geography-based theory of student demand that classifies consumers (prospective students) on behalf of consumers (colleges).  -->

Geomarkets and the Market Segment Model function together as an integrated market device that matches a socioeconomic hierarchy of places to a socioeconomic hierarchy of colleges.[^4] @RN4982 defined student market segments in ordinal terms -- local, in-state, regional, and national -- and found that student market segment was primarily a function of social class, particularly parental income and parental education. @RN4982 also conceived of colleges as an ordinal hierarchy. As in @RN4808, good matches are the key to realizing enrollment goals. In turn, @RN4982 argue that good matches are defined by homophily, the idea that actors who share common characteristics are likely to form connections with one another [@mcpherson2001birds; @RN4975]. @RN4982 conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges, @RN4982 "draw\[ing\] a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions" (p. 46). Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another. The takeaway is that like-colleges compete for like-students as defined by social class, with private selective colleges competing directly for students, charging the highest prices, and enrolling students with the highest socioeconomic status (p. 72):

[^4]: Whereas @RN5015 analyze devices that classify individuals, the Market Segment Model harkens back to the earlier era of geodemography, in which market research classified places for merchants based on the attributes of people who lived there [@RN4880; @RN4957]

> Students describe themselves socially simply by telling us the colleges and universities in which they are interested. The layering of collegiate competition is primarily a socioeconomic layering. The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.

## Research questions and expected results

Our overarching research question is, what is the relationship between college Board Geomarkets and high school recruiting visits of selective private colleges and public research universities? Colleges engage in enrollment management behaviors designed to realize enrollment goals. Whereas Geomarkets, the Market Segment Model, and EPS are market devices and/or components of market devices, the patterns of recruiting visits from colleges to high schools describe the performativity consequences of colleges using these market devices to realize enrollment goals.

<!-- SKETCHING THE ENROLLMENT ECONOMY
We sketch the enrollment economy for our analysis sample, consisting of 12 selective private liberal arts colleges, 14 selective private research universities, and 16 public research universities. Selective liberal arts colleges hews most closely to the characterization of @RN1549; their position in the zero-sum game for prestige substantially depends on the high school achievement of enrolled students. Primary revenue sources are tuition and donations from wealthy alumni. They primarily enroll rich students with a modicum of students from middle- and lower-class backgrounds [@RN4722]. Only a handful of colleges (e.g., Williams, Swarthmore) enjoy strong demand from students who are both affluent and high achieving. Outside the top ten or so, the priority becomes "affluent families \[who\] could absorb the College's full tuition and board costs with little, if any, financial aid" [@RN3519, p. 73]. Selective private research universities face a similar enrollment economy as liberal arts colleges, but are less dependent on students to the extent that they derive financial and reputional resources from faculty research. Following World War II, public research universities were less beholden to the enrollment economy due to federal research funding and stable state appropriations [@RN4517]. State appropriations became more volatile starting in the 1980s and declined substantially in the 2000s [@RN2271]. Public flagship universities responded by enrolling more nonresident students [@RN3753], who pay higher tuition prices and tend to be more affluent than resident students [@RN3685]. Therefore, state disinvestment makes public research universities beholden to the enrollment economy that governs selective private research universities, pursuing national recognition rankings and/or intercollegiate athletics as a means of generating demand from affluent households willing to pay the higher price of nonresident tuition.
-->


<!-- PARAGRAPH DESCRIBING DEPENDENCE BETWEEN COLLEGES AND HIGH SCHOOLS

Colleges depend on high schools to realize their enrollment goals. Although a college does not depend on any single high school to provide a substantial share of enrollment, the list of schools that enroll desirable students who are interested in attending is finite, thereby motivating "the recruiters' emphasis on feeder high schools" [@RN4982, p. 11]. High schools want to send their students to college. This goal is particularly important at expensive private schools and affluent public high schools, where placement into selective colleges functions as a key justification for both high tuition and high local housing prices. Drawing from @RN3519 and @RN4321, the relationship between college admissions officers and high school guidance counselors is one of mutual cooptation.^[The admissions officer conveys the special character of the college in hopes that the guidance counselor steers students to the college while the guidance counselor conveys the special value of the school's students, especially students deemed marginal by conventional measures of academic achievement.]  However, dyadic relationships are seldom symmetric [@RN960]. A college values a high school to the extent that they want to enroll students from that high school. A private college that depends mostly on tuition revenue may place minimal value on a high school that primarily enrolls low-income students. A high school values a college to the extent that they want some of their students to enroll in that college. The distribution of student achievement suggests that high schools value a range of colleges. Colleges below this range may be viewed with disdain while colleges above this range may be viewed as unattainable.

-->


Our first research question asks, to what extent are visit patterns consistent with recommendations from the Market Segment Model? The Market Segment Model is an algorithm for identifying desirable recruiting territories. For selective private -- and most public -- institutions in our sample, the Market Segment Model recommends focusing on affluent Geomarkets located in regions where many students are likely to consider a college like yours [@RN4982].

The US postsecondary education system is highly ordinal and so are student market segments. According to @RN4982, selective private colleges primarily depend on students from the regional and national market segments. Further, regional and national students tend to come from affluent, college-educated households. Affluent households are not scattered randomly across the country, but rather clump together in particular geographic communities. Like UK "league tables" [@RN5005], Geomarkets are a means of commensurating geographically disparate localities in terms of the number of students from each market segment. Imagine that the University of Arkansas has substantial regional but not national student demand and wants to enroll more out-of-state students. The Market Segment Model points to the affluent "TX 23 -- Collin & Rockwall" (near Dallas) and the "TX 15 -- Northwest Houston & Conroe School District" Geomarkets, which contain large numbers of regional and national students, but not the poorer "TX 19 -- City of Dallas" or "TX 17 -- City of Houston (East)," which contain relatively few regional and national students. We answer RQ1 using simple descriptive statistics. We predict most colleges in our sample will adhere to a regional or national (but not an in-state) recruiting strategy, depending on whether they have regional or national enrollment demand. We also predict that colleges will disproportionately visit high schools located in affluent Geomarkets.

RQ2 asks, to what extent do College Board Geomarkets (X) explain which high schools receive recruiting visits (Y)? We argue that Geomarkets explain college recruiting visits to high schools because they have become embedded in the organizational and technological infrastructures that generate them. These embeddings exemplify agencement, whereby Geomarkets transform from a mere classificatory device into a durable "socio-technical arrangement" of software, data fields, reports, and organizational roles, making Geomarkets a default unit for planning, staffing, and evaluation of recruitment. Through this agencement, Geomarkets may have performative consequences for recruiting behavior -- even after controlling for school- and neighborhood-level covariates -- because once a classification scheme is built into workflows, behavior will tend to align with it. We provide three examples of embedding Geomarkets into recruiting infrastructure.

First, although Geomarkets were initially designed to approximate the territories of admissions officers [@RN4982], over time they became an organizing principle through which admissions offices define recruiting territories and assign them to officers. On background, admissions officers and EM consultants told us that Geomarkets became part of the shorthand lexicon of the National Association of College Admissions Officers (NACAC). Admissions officers might introduce themselves by saying, "I recruit 'PA 2'," referring to the "Chester County, PA" Geomarket. We find evidence for this claim in the LinkedIn profiles of admissions officers. Examples include: "Personally managed schools in Indiana geomarkets 1, 2, 3, 4, 5, 6, 7, & 9" \[CITE\]; "Led recruitment efforts in geomarket MA06, the greater metro Boston area" [CITE]; and "Increased applications from the Long Island geomarket NY 16 - NY 21 for the 2021 admission cycle by 18.3%" [CITE]. Institutional materials also incorporate Geomarkets; a report from the Dartmouth College Admissions Ambassador Program lists the number of interviewers and applicants by Geomarket \[CITE\]. Anecdotal evidence suggests that admissions offices allocate relatively more admissions officers to affluent Geomarkets.^[For example, the University of Chicago assigns one recruiter to the "IL 9 - North Shore" Geomarket, which had a 2020 population of about 161,000, and one recruiter to the "IL 12 - South and Southwest Suburbs" Geomarket, which had a 2020 population of about 1.2 million [CITE]].


<!-- say something like, if no recruiter represents a Geomarket, then less likely to visit schools in that Geomarket

Assigning admissions recruiters to Geomarkets is a configuration of market devices and organizational routines that makes some recruiting actions more likely than others. 

files for all of these references can be found in here: /Users/jaquette/Documents/cb_geo/assets/images/evidence_of_geomarkets

cut text: ^[ Although Geomarkets are business-facing rather than consumer-facing devices, some admissions websites show concordance between Geomarkets and recruiting territories assigned to admissions officers. Examples include the Chicago-land recruiting territories of the University of Chicago \[CITE\], Lake Forest University, \[CITE\], and Saint Mary's College \[CITE\].] 

-->

Second, similar to the Black-Scholes model being written into spreadsheets used by options traders [@RN5067], Geomarkets and the logic of the Market Segment Model became the basis of the widely used EPS software, which encouraged colleges to identify promising Geomarkets and then select high schools to visit within them. EPS promotional material from  @eps_snapshot_2022 includes a screenshot of the user-interface navigation menu, which includes tabs for "Plan Travel," "Research High Schools," and "Competitive Analysis." This agencement integrates Geomarkets and the Market Segment Model logic into software-based workflows of territory assignment and travel planning, shaping the menu of possible recruiting visits and making some visits more likely than others. On background, EM professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories. EPS software often recommended visiting the same sets of affluent, high-achieving schools visited by other colleges. Some savvier, quantitatively adept offices, however, used EPS to visit schools that EPS recommended ignoring because there would be less competition for the good students who attended these schools. 

Third, Geomarkets have been integrated into Customer Relationship Management (CRM) and information systems used by colleges. Slate, the market-leading admissions CRM, allows institutions to assign admissions recruiters and prospective students to Geomarkets [CITE]. The LinkedIn profile of an admissions officer states, "Utilized Slate to create territory management tools for the recruitment team. Generated reports broken down by state and geomarket and sorted by high school to reflect applicant/admit/deposit data for past 3 cycles" [CITE]. Similarly, Oracle's PeopleSoft Campus Solutions student information system product integrates Geomarkets (referred to as "EPS Market Codes") and also EPS software [CITE]. These examples illustrate market devices as assemblages of heterogeneous components that build on one another. Integrating Geomarkets within third-party administrative infrastructuresmakes it easier for colleges to incorporate Geomarkets into recruiting workflows, including decisions about which high schools to visit.

<!--
Oracle software documentation reads \[CITE\]:

> EPS is a geographic and demographic data service offered annually by the College Board to Colleges and Universities. EPS provides information to subscribing institutions about competitors, feeder schools, and demographic strengths and weaknesses. EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States. Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
-->

RQ3 asks, to what extent does Geomarket popularity -- measured as the number of visits per school -- mediate the relationship between school composition and the probability of a receiving a visit? Previous research shows that having a high proportion of free-reduced lunch or Black/Hispanic students is negatively associated with receiving a visit [@RN4758]. Therefore, RQ3 investigates whether high- versus low-SES schools (and low versus high share of Black/Hispanic enrollment) benefit more from being located in a popular Geomarket. We hypothesize that high-SES and low Black/Hispanic enrollment schools benefit more from being located in a popular Geomarket. Our rationale draws from @RN4808, who highlight the role of "good matches" between ordinally classified products and ordinally classified customers. @RN5072 [pp. 20-21] offers an anecdote:

> One of the authors had a credit card transaction denied while trying to refill her car’s gas tank in a high-poverty Oakland neighborhood. The clerk indicated with resigned anger that the location of his gas station was the likely cause of the problem. Back in Berkeley a few minutes later, the card worked without a hitch.

Not coincidentally, Oakland is in Geomarket "CA 7 -- City of Oakland" whereas Berkeley is in Geomarket "CA 8 -- Alameda County excluding Oakland." @RN4808 [p. 17] write that "‘good matches’ are perceived as natural because they fit well with how things already are." An affluent school in an affluent Geomarket is a good match because it exemplifies the schools we expect to find. We predict that such schools benefit by receiving more visits than an affluent school in a less affluent Geomarket. By contrast, a poor or predominantly Black/Hispanic school in an affluent Geomarket does not match expectations and will not reap the same locational benefits.



# Methods

<!-- 

Data → Variables → Analytic Strategy

Data

Recruiting visit data (collection, sample of 42 colleges)
Secondary data sources


Variables

Outcome
Geomarket measures (key independent variables)
School and neighborhood covariates


Analytic Strategy

RQ1: Descriptive statistics
RQ2: Fixed effects regression
RQ3: Interaction models

-->

## Data and Sample

**Sample of colleges**. This manuscript draws from a data collection that tracked off-campus recruiting visits made during the 2017 calendar year by a convenience sample of colleges. The data collection sample consisted of: all public research-extensive universities as defined by the 2000 Carnegie Classification (N=102), all private universities in the top 100 of the 20XX U.S. News and World Report National Universities rankings (N=57), and all private colleges in the top 50 of the 20XX U.S. News and World Report Liberal Arts Colleges rankings (N=47). The rationale for different criteria for inclusion across different university types is because these criteria reflected the interest of a foundation that funded the data collection. 

For universities in our data collection sample, we investigated their admissions website for pages that posted their upcoming off-campus recruiting visits. For institutions that posted such pages, we scraped the pages once per week throughout 2017.^[Web-scraping was conducted using the Python programming language using the _requests_ library<!--[CRYSTAL -- STATE WHICH PYTHON PACKAGE? AND/OR SHOW LINK TO A SAMPLE SCRIPT?] TODO: Added the package used but not too sure if there's a simple way to show example... these (https://github.com/ksalazar3/recruitment-scraping/tree/master/universities) are univs w/ custom scripts but they're all passed to this main script (https://github.com/ksalazar3/recruitment-scraping/blob/master/scraper.py) that actually calls the requests library. Also, this GitHub repo is currently private.-->.] Many universities in our data collection sample only posted certain kinds of events (e.g., hotel receptions and national college fairs) but not others (e.g., day-time visits to high schools). These institutions were excluded from the analysis sample. The final analysis sample for this article consists of 16 public research universities, 14 private selective research universities, and 12 private seletive liberal arts colleges that we believed posted all varieties of off-campus recruiting events on their admissions website.^[For public universities in our analysis sample, we assessed the quality of the web-scraped data by asking the enrollment/admissions office for these data. If a university denied or ignored our request, we issued a formal public records request. Of the 16 public universities analyzed in this study, we received requested data from 11 universities. For these 11 universities, web-scraped data was very similar to publicly requested data. For universities that sent us data, the analyses below use "requested" rather than "scraped" data because requested data tended to have more events. Broad patterns were similar across requested data versus scraped data, and results based on scraped data are available upon request. @RN4758 describe data quality checks in greater detail.]

<!-- SHOULD WE INCLUDE SOMETHING LIKE THIS?

Our analysis sample of universities should be considered a convenience sample. Table \@ref(tab:sample-characteristics) shows selected characteristics for the universities in our analysis sample. Table \@ref(tab:sample-population-characteristics) shows that our analysis samples of public and private universities appear broadly similar to the populations they were drawn from. 

INSERT TABLES \@ref(tab:sample-characteristics) AND \@ref(tab:sample-population-characteristics) ABOUT HERE

-->



**Sample of high schools**. The analysis sample of high schools consists of all public and private high schools that met the following criteria in YYYY [what year]: enrolls at least 10 12th grade students: is not a special education school; is not primarily a virtual school; and is located in the 50 U.S. states, the District of Columbia, or land regulated by the Bureau of Indian Affairs. The final analysis sample included 20,328 public high schools and 4,126 private high schools. The public school sample consisted of 18,110 regular, 216 career/technical, and 2,002 alternative education schools. The private school sample consisted of 3,766 regular, 11 career/technical, 147 alternative education, and 202 special program emphasis [RECHECK/REVISE NUMBERS].

Analyses utilized several secondary data sources. The data on university characteristics come from the 2017 Integrated Postsecondary Education Data System (IPEDS) and 2020 Best Colleges ranking by U.S. News and World Report. Public high school measures are derived from the 2017-18 NCES Common Core of Data (CCD) and Niche 2020 Best Public High Schools ranking. Private high school measures are derived from 2017-19 NCES Private School Universe Survey (PSS) and Niche 2020 Best Private High Schools ranking.^[For both public and private high schools, we also used NCES data from previous years when the 2017-18 data was not available for a small number of schools.] 

NEIGHBORHOOD CHARACTERISTICS CREATED USING ACS DATA MEASURED AT ZIP CODE OR TRACT LEVEL

## Variables

The dependent variable, $Y_{i,c}$ is a dichtomous indicator of whether a high school $i$ received one or more recruiting visits in 2017 from college $c$ in our sample. 

The independent variable of interest is College Board Geomarket. We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by The College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products [@collegeboard_connections_geographies]. Specifically, we georeferenced the College Board’s Geomarket maps by overlaying PDF images onto spatial coordinates aligned with census tract boundaries. We then extracted vector data by manually tracing the Geomarket boundaries and generating individual shapefiles for each Geomarket. For instance, @fig-geomarket-methods shows The College Board's map of Geomarkets in the Chicago metropolitan area on the left and our created shapefiles on the right. This approach to constructing Geomarkets aligned precisely with 2020 census tract boundaries from ACS 5-year estimates; that is, each census tract corresponded uniquely to a Geomarket generated from the College Board images. For example, @fig-geomarket-misalign zooms into the northern Geomarkets of the Chicago metropolitan area from @fig-geomarket-methods. The center map showcases how our created Geomarket shapefiles (delineated in purple) align with the spatial coordinates of 2020 census tract boundaries (delineated in grey). 

## Analytic strategy

# Results

## Research Question 1

Order of results

- freshman enrollment by state
- recruiting visits by student market segment
- recruiting visits by EPS region
- gini coefficient
- top 30 visited geomarkets
  - might create small multiples
    - 3 by 3?: 
      - columns: private universities, liberal arts colleges, public research, 
      - rows: in-state, regional visits, national visits

RQ1 asks, to what extent are visit patterns consistent with recommendations by the Market Segment Model? To provide context, Figure X shows the geographic origin of 2016 Freshman enrollment for colleges in our sample. On the left panel, private colleges have a higher share of out-of-state enrollment, but several public research universities -- e.g., Pitt, University of Alabama, University of South Carolina, University of Arkansas, and CU-Boulder -- rely heavily on nonresident enrollment. The right panel categorizes domestic freshman enrollment by EPS region. Colleges are sorted by type -- private research, liberal arts, public research -- and then EPS region. Applying the Market Segment Model to the case of enrollment, most private colleges have a "regional" enrollment strategy, where the majority of freshman are from the same EPS region as the college. Consistent with ,

For all publics and most privates, the majority of freshman hail from the same EPS region 





## Research Question 2

## Research Question 3



# Discussion

<!-- discussion should establish concept infrastructural collusion -- or at least discuss the idea. and should give the COFHE paragraph, which highlights explicit collusion. 

o	P. 45: “if students are aware of the social patterning that so often underlies their college choices, they seldom talk about it except to note, “My friends and I are considering the same kind of schools.” Students talk in shorthand phrases about different types of colleges, classifying them on the basis of similarly perceived offerings and settings. Most admissions officers, even most presidents and provosts, intuitively understand that students think not so much of individual institutions as of groups of institutions In the words of one veteran dean of admissions, “We are known primarily by the company we keep”
o	“For this reason, colleges actually encourage competition with one another to define their places in the market. Members of the Consortium on Financing Higher Education (COFHE) – a group of thirty highly selective, high-cost, private institutions – regularly exchange detailed and confidential information about their applicant pools, admissions strategies, and financial aid programs. What every envious admissions officer outside the Consortium understands is that these thirty institutions are strengthened by their collectivity, though they are in fact one another’s principal competitors. The dean of admissions at one COFHE institution knows that there is greater chance of interesting a student who is already considering another COFHE institution than a student who is not. The same kind of forces help define collective markets for state colleges, women’s colleges, Catholic or other church-related institutions, and even diverse urban institutions that share intense athletic rivalries.”



BELONGS SOMEWHERE IN DISCUSSION
Economics conceives of matching mechanisms, such as the SAT, as contributing to an "efficient outcome" [@RN2247] whereby high-aptitude students are matched to colleges with sufficient resources for these students to realize their potential. Although this argument maybe plausible for for Ivy+ institutions [@RN4935], prior research suggests that the majority of selective private colleges and public research universities prioritize "affluent families \[who\] could absorb the College's full tuition and board costs with little, if any, financial aid" [@RN3519, p. 73].
-->

<!-- CUT TEXT ON UTILITY OF MARKET DEVICES FOR EMPIRICAL RESEARCH. MOVE THIS TO DISCUSSION. 

The market devices literature offers great utility for the sociology of education subfield. The market devices literature examines how techniques, theories, culture, and agency interract in ways that make economic action possible [@RN4777; @RN5087]. Market devices are built upon other market devices and products are market devices that are composed of many market devices. This perspective legitimizes scholarly attention to individual components that are foundational to products and to markets [@RN5071], facilitating scholarship that sacrifices breadth for depth. For example, rather than analyzing insurance products as a whole, @RN5070 focuses on the statistical models that assign people to different risk pools. Moreover, the market devices literature legitimizes scholarly attention to the technical details of how market devices are constructed because technical details have consequences [@RN5067; @RN5078]. The flexibility of scope offered by the market devices literature is poignant for the sociology of education as third-party products increasingly perform core functions on behalf of schools and colleges. For scholarship that traces how a set of theories or techniques become embedded in one or more products, the market devices literature answers the essential question, ‘my study is a case of what?’ The market devices literature is situated within the economic sociology subfield. After some debate \[CITE\], it became part of the orthodoxy. Sociology of education draws heavily from organizational sociology [@RN5097] but less so from economic sociology and, to our knowledge, has not substantively utilized the market devices literature as a foundation for empirical scholarship.

-->

```{=html}
<!-- 

> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting\ldots We have begun with this celebration of storytelling\ldots \[because\] we believe that the intuitions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research is based upon listening carefully to what admissions officers have to say (pp. 9-10).

The Market Segment Model is the product of practioner-derived market research. In the 2000s, College Board created the *Segment Analysis Service* product [@RN4565] as an application of geodemographic market research logics -- associated with mail-order catalogs -- to the case of college choice [@RN4957]. By contrast, @RN4982 conceived of the Market Segment Model as an effort to "capture and quantify" (p. 11) the knowledge of admissions officers:
-->
```

# References

\singlespacing
\begingroup
\small

::: {#refs}
:::

\clearpage
\newgeometry{margin=0.1in}

```{=latex}
\begin{landscape}
```

```{r fig-fresh-enroll, echo=FALSE, out.width="100%", fig.cap="Geographic origin of fall 2016 freshman enrollment"}
knitr::include_graphics("../results/fall_2016_freshman_enroll_combined_plots.pdf")
```



```{r fig-recruiting-heatmap-market-segment, echo=FALSE, out.width="90%", fig.cap="Recruiting visits by market segment of Geomarket"}
knitr::include_graphics("../results/recruiting_heatmap.pdf")
```

```{r fig-recruiting-heatmap-eps-region, echo=FALSE, out.width="90%", fig.cap="Recruiting visits by EPS region"}
knitr::include_graphics("../results/recruiting_heatmap_eps_region.pdf")
```

```{r}
#| label: tbl-top30-pub
#| tbl-cap: 'Top 30 Geomarkets ranked by visits per school (public school visits only)'

top30_geo_pub_sch <- readRDS('../results/top30_geo_pub_sch.RDS')
top30_geo_pub_sch %>% 
  kbl(
    booktabs = TRUE,
    digits   = 1
  ) %>%
  kable_classic(full_width = FALSE)
```

```{r}
#| label: tbl-top30-priv
#| tbl-cap: 'Top 30 Geomarkets ranked by visits per school (private school visits only)'

top30_geo_priv_sch <- readRDS('../results/top30_geo_priv_sch.RDS')
top30_geo_priv_sch %>% 
  kbl(
    booktabs = TRUE,
    digits   = 1
  ) %>%
  kable_classic(full_width = FALSE)
```

```{=latex}
\end{landscape}
```

```{r fig-concentration, echo=FALSE, out.width="80%", fig.cap="Lorenz-style concentration curves of recruiting visits by Geomarket affluence"}
knitr::include_graphics("../results/combined_concentration_graph.pdf")
```

```{=latex}
\begin{landscape}
```

```{r}
#| label: tbl-model-fit-stats
#| tbl-cap: 'Model Fit Statistics, fixed effects regression of visit to high school i by college j, separate models for all HS, public HS, private HS'

tab_df_all_ij <- readRDS('../results/reg_tab_all_ij.RDS')
tab_df_pub_ij <- readRDS('../results/tab_df_pub_ij.RDS')
tab_df_priv_ij <- readRDS('../results/tab_df_priv_ij.RDS')

bind_rows(
  tab_df_all_ij %>% tail(4),
  tab_df_pub_ij %>% tail(4),
  tab_df_priv_ij %>% tail(4)
) %>% 
  mutate(
    across(
      .cols = -term,
      .fns = ~ ifelse(
        term == "N",
        format(
          round(as.numeric(.), 0),
          big.mark    = ",",
          scientific  = FALSE,
          trim        = TRUE
        ),
        .
      )
    )
  ) %>% 
  kbl(
    booktabs = TRUE,
    latex_options = c('scale_down')
  ) %>% 
  pack_rows('All High Schools', 1, 4) %>% 
  pack_rows('Public High Schools Only', 5, 8) %>% 
  pack_rows('Private High Schools Only', 9, 12)
```

```{=latex}
\end{landscape}
```

```{r}
#| label: tbl-regression-interaction
#| tbl-cap: 'Fixed Effects Regression Model with Interaction between Geomarket Popularity and School SES/Racial Composition'

#model <- readRDS('../results/model.RDS')
#model_interact_frl <- readRDS('../results/model_interact_frl.RDS')
#model_interact_blhisp <- readRDS('../results/model_interact_blhisp.RDS')
model <- readRDS('../../cb_geomarket_shape/modelout/model.RDS')
model_interact_frl <- readRDS('../../cb_geomarket_shape/modelout/model_interact_frl.RDS')
model_interact_blhisp <- readRDS('../../cb_geomarket_shape/modelout/model_interact_blhisp.RDS')


stars_cutoffs <- c(
  '***' = 0.001,
  '**' = 0.01,
  '*' = 0.05
)

add_stars <- Vectorize(function(p) {
  for (s in names(stars_cutoffs)) {
    if (p < stars_cutoffs[s]) {
      return(s)
    }
  }
  return('')
})

get_stats <- function(m) {
  data.frame(
    variable = c('N', 'Adj. R²', 'Within R²', 'RMSE'),
    Estimate = c(
      format(m$nobs, big.mark = ',', scientific = F),
      sprintf('%.3f', r2(m)[['ar2']]),
      sprintf('%.3f', r2(m)[['wr2']]),
      sprintf('%.3f', sqrt(mean(resid(m)^2)))
    )
  )
}

stats_df <- get_stats(model) %>%
  left_join(get_stats(model_interact_frl), by = 'variable') %>%
  left_join(get_stats(model_interact_blhisp), by = 'variable')

model_df <- as.data.frame(model$coeftable) %>% rownames_to_column('variable') %>% mutate(Estimate = str_c(sprintf('%.3f', Estimate), add_stars(`Pr(>|t|)`))) %>% 
  full_join(as.data.frame(model_interact_frl$coeftable) %>% rownames_to_column('variable') %>% mutate(Estimate = str_c(sprintf('%.3f', Estimate), add_stars(`Pr(>|t|)`))), by = 'variable') %>% 
  full_join(as.data.frame(model_interact_blhisp$coeftable) %>% rownames_to_column('variable')%>% mutate(Estimate = str_c(sprintf('%.3f', Estimate), add_stars(`Pr(>|t|)`))), by = 'variable') %>% 
  filter(str_detect(variable, 'peps_n_vis01_per_sch_all|hs_pct_free_reduced_lunch_decile|hs_pct_bl_hisp_nat_decile')) %>% 
  arrange(desc(str_detect(variable, 'hs_pct_free_reduced_lunch_decileD\\d+$'))) %>%
  select(-contains('t value'), -contains('Pr')) %>% 
  mutate(across(everything(), ~ {
    if(is.numeric(.)) {
      # Numeric: round + format to 3 decimals as string
      s <- sprintf("%.3f", .)
      s[is.na(.)] <- ""  # replace NAs with empty string
      s
    } else {
      # Non-numeric: leave as-is, replace NA with ""
      s <- as.character(.)
      s[is.na(s)] <- ""
      s
    }
  }))

model_df %>% 
  bind_rows(stats_df) %>% 
  kbl(
    booktabs = TRUE,
    col.names = c('', rep(c('Estimate', 'Std Error'), 3)),
    linesep = c(rep(c(rep('', 8), '\\addlinespace'), 2), '\\addlinespace', rep(c(rep('', 8), '\\addlinespace'), 2), rep('', nrow(stats_df)))
  ) %>%
  kable_styling(
    latex_options = c('scale_down')
  ) %>% 
  add_header_above(c(' ' = 1, 'Model' = 2, 'Interaction Free-Reduced Lunch' = 2, 'Interaction Black/Hisp' = 2)) %>%
  row_spec(nrow(model_df), extra_latex_after = paste0("\\addlinespace\\cline{2-", length(model_df), "}")) %>%
  footnote(
    general_title = '',
    general = str_c('Significance codes: ', str_c(names(stars_cutoffs), ' p<', stars_cutoffs, collapse = ', ')),
    threeparttable = TRUE,
    escape = FALSE
  )
```

\restoregeometry
\clearpage

# Appendix

\setcounter{table}{0} 
\renewcommand{\thetable}{A\arabic{table}}
\clearpage
\newgeometry{margin=0.1in}
```{=latex}
\begin{landscape}
```

```{r, eval = T}
#| label: tbl-regression-pub
#| tbl-cap: 'Fixed effects regression models of visit to high school i by college j (public high schools only)'

tab_df_pub_ij <- readRDS('../results/tab_df_pub_ij.RDS')
display_regression_table(tab_df_pub_ij, font_size = 9)
```

```{r, include = FALSE}
#| label: tbl-regression
#| tbl-cap: 'Fixed effects regression models of visit to high school i by college j (all high schools)'

tab_df_all_ij <- readRDS('../results/reg_tab_all_ij.RDS')
display_regression_table(tab_df_all_ij)
```

```{r, eval = T, include = FALSE}
#| label: tbl-regression-priv
#| tbl-cap: 'Fixed effects regression models of visit to high school i by college j (private high schools only)'

tab_df_priv_ij <- readRDS('../results/tab_df_priv_ij.RDS')
display_regression_table(tab_df_priv_ij, font_size = 9)
```

\clearpage

```{r}
#| label: tbl-adjr2-pub
#| tbl-cap: 'Model fit statistics, fixed effects regression models of visits to high school i, separate models models by college j (public high schools only) '

adjr2_pub_with_meta <- readRDS('../results/adjr2_pub_with_meta.RDS') %>% select(-univ_classification, -withinR2_pub_cov_eps_fe)

adjr2_pub_with_meta %>% 
  kbl(
    booktabs = TRUE,
    digits   = 2,
    col.names = c(
      'University', 'USNWR Rank', '# PubHS Visits',
      'AdjR2 State Only', 'AdjR2 EPS FE Only', 'AdjR2 Cov Only', 'AdjR2 Cov State', 'AdjR2 Cov EPS FE'
    )
  ) %>%
  kable_styling(
    font_size = 9,
    latex_options = c('scale_down')
  ) %>% 
  pack_rows('Private Liberal Arts', 1, 12) %>% 
  pack_rows('Private Research', 13, 26) %>% 
  pack_rows('Public Research', 27, 42)
```

```{r, include = FALSE}
#| label: tbl-adjr2
#| tbl-cap: 'Model fit statistics, fixed effects regression models of visits to high school i, separate models models by college j (all high schools) '

adjr2_all_with_meta <- readRDS('../results/adjr2_all_with_meta.RDS') %>% select(-withinR2_cov_eps_fe, -univ_classification)

adjr2_all_with_meta %>% 
  kbl(
    booktabs = TRUE,
    digits   = 2,
    col.names = c(
      'University', 'USNWR Rank', '# Visits', '# PubHS Visits', '# PrivHS Visits',
      'AdjR2 State Only', 'AdjR2 EPS FE Only', 'AdjR2 Cov Only', 'AdjR2 Cov State', 'AdjR2 Cov EPS FE'
    )
  ) %>%
  kable_styling(
    latex_options = c('scale_down')
  ) %>% 
  pack_rows('Private Liberal Arts', 1, 12) %>% 
  pack_rows('Private Research', 13, 26) %>% 
  pack_rows('Public Research', 27, 42)
```

```{r, include = FALSE}
#| label: tbl-adjr2-priv
#| tbl-cap: 'TBD (Private High Schools Only)'

adjr2_priv_with_meta <- readRDS('../results/adjr2_priv_with_meta.RDS') %>% select(-univ_classification, -withinR2_priv_cov_eps_fe)

adjr2_priv_with_meta %>% 
  kbl(
    booktabs = TRUE,
    digits   = 2,
    col.names = c(
      'University', 'USNWR Rank', '# PrivHS Visits',
      'AdjR2 State Only', 'AdjR2 EPS FE Only', 'AdjR2 Cov Only', 'AdjR2 Cov State', 'AdjR2 Cov EPS FE'
    )
  ) %>%
  kable_styling(
    font_size = 9,
    latex_options = c('scale_down')
  ) %>% 
  pack_rows('Private Liberal Arts', 1, 12) %>% 
  pack_rows('Private Research', 13, 26) %>% 
  pack_rows('Public Research', 27, 42)
```

```{=latex}
\end{landscape}
```

\restoregeometry
\clearpage
