---
title: "The Road Not Taken: College Board Geomarkets and Recruiting Visits by Selective Colleges"
# author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)", "Crystal Han (Stanford University)"]
bibliography: ./../../cb_geo/assets/bib/cb_geomarket_bib.bib
csl: ./../../cb_geo/assets/bib/chicago-author-date.csl
format:
  #docx:
  pdf:
    documentclass: article
    #papersize: letter
    number-sections: true
    number-depth: 2
    #toc: FALSE
    #pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #keep_tex: true
editor: visual
geometry:
  - margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
library(knitr)
#library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)

#hey ho This many words: {{< words-total >}}
```

\pagenumbering{roman}
\newpage

\pagenumbering{gobble}
\pagenumbering{arabic}


# section title

Overarching RQ

- What is the relationship between College Board Geomarkets (X) and high school recruiting visit patterns?

Concrete RQs

- To what extent do college board Geomarkets explain which high schools receive recruiting visits from public research and selective private PSIs? 
- To what extent are visit patterns consistent with recommendations from the Market Segment?
- If and to what extent does geomarket popularity [Z] mediate the relationship between school composition [free reduced lunch; share Black/Hispanic] and probability of a high school getting a visit [Y]?

@fig-fresh-enroll here

# References

\singlespacing
\begingroup
\small


::: {#refs}
:::

# Appendix

\clearpage
\newgeometry{margin=0.1in}

```{r fig-fresh-enroll, echo=FALSE, out.width="100%", fig.cap="Geographic origin of fall 2016 freshman enrollment"}
knitr::include_graphics("../results/fall_2016_freshman_enroll_combined_plots.pdf")
```

```{=latex}
\begin{landscape}
```

```{r fig-recruiting-heatmap-market-segment, echo=FALSE, out.width="90%", fig.cap="Recruiting visits by market segment of Geomarket"}
knitr::include_graphics("../results/recruiting_heatmap.pdf")
```

```{r fig-recruiting-heatmap-eps-region, echo=FALSE, out.width="100%", fig.cap="Recruiting visits by EPS region"}
knitr::include_graphics("../results/recruiting_heatmap_eps_region.pdf")
```

```{=latex}
\input{"../results/top30_geo_pub_sch.tex"}
```


```{=latex}
\input{"../results/top30_geo_priv_sch.tex"}
```

<!--
```{=latex}
\input{"../results/top30_geo_pub_sch_national.tex"}
```

```{=latex}
\input{"../results/top30_geo_priv_sch_national.tex"}
```
-->

```{=latex}
\end{landscape}
```

\restoregeometry
\clearpage
